---
title: "Validation summary"
output: html_notebook
---

```{r echo=F}
library(ggplot2)
library(knitr)
library(gridExtra)
```

# Validation results for Bayesian GGMs calculated on KORA and LOLIPOP cohort

This document is a (graphical) summary of the results from the validation output.
It uses the results collected in $results/validation.summary.txt$.

## Mediation results

```{r echo=F}
# load the large result table for all loci
# TODO we currently remove the results for the GO enrichment
tab <- read.table("results/current/validation.summary.txt", header=T, sep="\t")[,1:15]

# create a mediation specific plot
mediation <- tab[,c(1,2,7,8,9)]
# for now we ignore results where we didn't have SNP genes at all
mediation <- mediation[!is.infinite(mediation$log10_med), ,drop=F]

# plot the mediation results
toplot <- mediation[order(mediation$mediation_selected),]
toplot$sentinel <- factor(toplot$sentinel, levels=unique(toplot$sentinel))
toplot <- melt(toplot, measure.vars=c(3,4), value.name = "pval")

p <- ggplot(data=toplot, aes(x=sentinel, y=-log10(pval), fill=variable)) +
  geom_bar(stat="identity", position="dodge") + 
  facet_grid(cohort ~ .) +
  theme(axis.text.x = 
          element_text(angle = 90, hjust = 1, vjust=0))
p
```

Above the mediation results on the different cohorts for all hotspots 
in which at least one SNP gene was extracted by our algorithm are shown.
The *mediation_selected* group shows the significance of the mediation analysis of 
those selected genes, the *mediation* group shows the significance of all NOT selected genes. 
[Several hotspots are currently not being shown, since we had either problems with 
the GGM calculation or the no SNP genes were selected.]


```{r echo=F}
# in addition we plot the log fold-change
# sort by log_fc
toplot <- mediation[order(mediation$log10_med, decreasing = T),]
toplot$sentinel <- factor(toplot$sentinel, levels=unique(toplot$sentinel))

toplot <- melt(toplot, measure.vars=5, value.name="log_fc")
toplot$favored <- factor(sign(toplot$log_fc),labels = c("not selected", "selected"))
p <- ggplot(toplot, aes(x=sentinel, y=log_fc, fill=favored)) +
  geom_bar(stat="identity") + 
  facet_grid(cohort~.) +
  theme(axis.text.x = 
          element_text(angle = 90, hjust = 1, vjust=0)) +
  ylab("log10(all / selected)")
p

perc <- table(toplot$favored)
perc <- perc["selected"] / sum(perc)

```
This figure shows a different summary of the mediation results. Shown is the log10 fold change
of the mediation p-value for the not selected SNP genes ($p_n$) over the p-value for selected
genes ($p_s$). The red bars (*not selected*) indicate fold changes where $p_n$ is lower 
than the corresponding $p_s$, whereas the blue bars indicate the opposite. Overall, 
**`r format(perc*100,digits=4)`\%** of all fold changes show negative fold changes 
(i.e. $p_s$ being smaller than their respective $p_n$s). Currently, we select the
minimal p-value obtained from the genes not selected via a GGM and the maximal p-value
from the selected genes which is a rather conservative estimate. We could think about
changing the way of summarizing the p-values...

## Gene-Gene validation 

The gene-gene links were validated using external expression data from GEO (ARCHS4),
Geuvadis as well as the data from either LOLIPOP or KORA, depending on which cohort 
the bGGM was calculated.

To assess how well our approach recovers co-expression of genes in the independent
datasets, we calculate all gene-vs-gene (gvg) correlations in those data and create a
confusion matrix as follows:

```{r results='asis', echo=F }
temp <- matrix(c("W","X","Y","Z"),nrow=2,ncol=2)
colnames(temp) <- c("In GGM", "Not in GGM")
rownames(temp) <- c("Correlation sign.","Correlation not sign.")
kable(temp,align = c("c","c"))
```
We deem a correlation between two genes to be significant if 1) $p-value < 0.01$
and 2) $\rho > 0.3$ ($\rho$ being the Pearson Correlation Coefficient).

We test whether we can reject $H0$ (i.e. whether we detect a correlation regardless of it 
being significant in the independent data) using a Fisher's exact test.
Shown below are the p-values for those tests over all loci on the different datasets.
The black lines indicate a significance level of $0.05$.
```{r echo=F}
# get the relevant data (pvalues on the different datasets)
gg <- tab[,c(1,2,13,14,15)]
colnames(gg) <- c("sentinel", "cohort", "Geuvadis", "GEO", "LOLIPOP_KORA")
#gg <- gg[order(gg$LOLIPOP_KORA),]
#gg$sentinel <- factor(gg$sentinel, levels=unique(gg$sentinel))

# plot the data
gg <- melt(gg,measure.vars=c(3,4,5), id.vars=c(1,2))
gg$value <- -log10(gg$value)
p <- ggplot(gg, aes(x=reorder(sentinel,value), y=value, fill=variable)) + 
  facet_grid(cohort~variable) +
  geom_bar(stat = "identity", position="dodge") + 
  theme(axis.text.x = element_text(vjust=1, angle = 90)) + 
  xlab("sentinel") +
  geom_hline(yintercept = -log10(0.05))

p

#create list of plots
#p <- lapply(split(gg,gg$variable), function(x){
  
#  x$sentinel <- factor(x$sentinel, levels=unique(x$sentinel[order(-log10(x$value))]))

  # create the plot
#  p <- ggplot(x, aes(x = sentinel, y = -log10(value), fill = variable)) +
#    geom_bar(stat = "identity") +
#    facet_grid(cohort~.) +
#    theme(legend.position="none", axis.text.x = element_text(vjust=1, angle = 90)) + 
#    geom_hline(yintercept = -log10(0.05)) +
#    labs(title=unique(x$variable))
#})

#do.call(grid.arrange,(c(p, nrow=3)))
```


