---
title: "Replication performance of prior based methods (noise evaluation)."
author: "Johann Hawe"
date: 2021-03-30
output:
  html_document:
    df_print: paged

---

```{r setup, echo=FALSE, warning=FALSE, message = FALSE}
library(tidyverse)
library(graph)
source("scripts/validation_methods.R")

knitr::opts_chunk$set(echo=FALSE)
threads <- 4

base_dir <- "results/current/biogrid_stringent/fits_tfa/_rerun/replication_prior_noise/"
```

In this document we investigate the effect of noise in prior information on replication performance.
To successfully perform the analysis, we expect that the two snakemake rules `all_ggm` and `all_ggm_with_prior_noise` have been run and results are available under `r base_dir`.

```{r load_data}
fit_files <- list.files(base_dir, "*.rds", full.names = T)
```

We are currently looking at the results for `r length(fits)` loci.
Let's have a look at the overall replication 'performance' in terms of MCC.

```{r get_replication_performance_mcc}
models <- c("bdgraph", "bdgraph_no_priors", "glasso", "glasso_no_priors",
            "genenet", "genie3", "irafnet")

result_table <- parallel::mclapply(fit_files, function(file) {
  
  file_name <- gsub(".rds", "", basename(file))
  sentinel <- gsub("_meqtl|_eqtlgen", "", file_name)
  locus_type <- gsub("rs.*_", "", file_name)
  
  fits <- readRDS(file)
  
  lapply(names(fits), function(noise_level) {
    
    fits_at_noise_level <- fits[[noise_level]]
    
    fits_kora <- fits_at_noise_level[["kora"]]
    fits_lolipop <- fits_at_noise_level[["lolipop"]]
    
    lapply(models, function(model) {
      fit_kora <- fits_kora[[model]]
      fit_lolipop <- fits_lolipop[[model]]
      
      replication <- get_graph_replication_f1_mcc(fit_kora, fit_lolipop)
      # replication null means that there was no overlap between graphs in terms
      # of common nodes
      if(is.null(replication)) {
        replication <- list(MCC = 0, F1 = 0)
      } 
      
      tibble(sentinel = sentinel, locus_type = locus_type,
             model = model, noise_level = gsub("noise_level_", "", noise_level),
             mcc = replication$MCC, f1 = replication$F1)
      
    }) %>% bind_rows()
    
  }) %>% bind_rows()
  
}, mc.cores = threads) %>% bind_rows() %>%
  mutate(model = gsub("^glasso$", "glasso (priors)", model)) %>%
  mutate(model = gsub("^bdgraph$", "bdgraph (priors)", model)) %>%
  mutate(model = gsub("_no_priors", "", model))
  
result_table
```

We got replication performances for `r nrow(result_table)` combinations of loci, models and noise levels. Now we can plot those results.

```{r plot_replication_performance}
# define colors
paired <- RColorBrewer::brewer.pal(4, "Paired")
names(paired) <- c("glasso", "glasso (priors)", "bdgraph", "bdgraph (priors)")
unpaired <- RColorBrewer::brewer.pal(7, "Dark2")[c(2,3,7)]
names(unpaired) <- c("irafnet", "genie3", "genenet")
graph_cols <- c(paired, unpaired)

ggplot(result_table, 
       aes(x=noise_level, y=mcc, color=reorder(model, -mcc))) + 
  geom_boxplot() + 
   stat_summary(
    fun.y = median,
    geom = "smooth",
    position = position_dodge2(width = 0.9),
    aes(group = reorder(model, -mcc)),
    lwd = 0.8
  ) + 
  scale_color_manual(values = graph_cols) + 
  labs(color = "method", y="MCC",
       x = "Level of prior noise")
  
```

```{r session_info}
sessionInfo()
```

