---
title: "Additional simulation analysis based on reviewer comments"
output: 
 html_notebook:
    toc: TRUE
---

```{r setup}
library(dplyr)
library(readr)
library(grid)
library(ggplot2)
library(cowplot)
library(RColorBrewer)
library(ggpubr)
library(scales)
library(reshape2)
library(circlize)
library(ggbio)
library(xtable)

# get the hg19 chromosome definitions
library(BSgenome.Hsapiens.UCSC.hg19)
hg19info <- seqinfo(BSgenome.Hsapiens.UCSC.hg19)

# set up theme and colors
theme_set(theme_cowplot() + background_grid())
theme_update(legend.text = element_text(size=11), 
             legend.title=element_text(size=12),
             axis.text.x = element_text(size=10),
             axis.text.y = element_text(size=10))

# manually define the colors for the different methods (to get paired ones for
# prior/non prior versions)
paired <- brewer.pal(4, "Paired")
names(paired) <- c("glasso", "glasso (priors)", "bdgraph", "bdgraph (priors)")
#paired <- c(paired, "#66D35F")
#names(paired) <- c("glasso", "glasso (priors)", 
#                   "bdgraph (empty)", "bdgraph (priors)", "bdgraph (full)")
unpaired <- brewer.pal(7, "Dark2")[c(2,3,7)]
names(unpaired) <- c("irafnet", "genie3", "genenet")
graph_cols <- c(paired, unpaired)

sfb_graphs <- scale_fill_manual(values=graph_cols)
#sfb_graphs <- scale_fill_brewer(palette="Set2")
#scb_graphs <- scale_color_brewer(palette="Set2")
scb_graphs <- scale_color_manual(values=graph_cols)
sfb_binary <- scale_fill_brewer(palette = "Accent")
scb_binary <- scale_color_brewer(palette = "Accent")
scb_priors <- scale_color_brewer(palette = "Dark2")

bgm <- background_grid(major = "xy")
group_cols <- brewer.pal("Set2", n=3)
COLORS <- list(MEQTL = group_cols[1],
               EQTL = group_cols[2])

```

```{r}

```

```{r minimal_subset}
# minimal subset size, prior noise progression
res <- read_tsv("temp") %>%
  mutate(R = paste0("R=", rdegree))

# create nicer method names
tab <- res %>% 
  mutate(comparison = gsub("bdgraph$", "bdgraph (priors)", comparison),
         comparison = gsub("glasso$", "glasso (priors)", comparison),
         comparison = gsub("bdgraph_no_priors$", "bdgraph (empty)", comparison),
         comparison = gsub("bdgraph_no_priors_full$", "bdgraph (full)", comparison),
         comparison = gsub("glasso_no_priors","glasso", comparison)) %>%
  filter(comparison != "bdgraph (full)") %>%
  dplyr::rename(method=comparison) %>%
  mutate(method = gsub("bdgraph \\(empty\\)", "bdgraph", method)) %>%
  mutate(method = factor(
    method,
    ordered = T,
    levels = c(
      "bdgraph",
      "bdgraph (priors)",
      "glasso",
      "glasso (priors)",
      "irafnet",
      "genie3",
      "genenet"
    )
  ))

# get the MCC plot
simulation_mcc <- ggplot(tab,
                         aes(y = MCC,
                             x = R,
                             color = method)) +
  stat_boxplot(geom = "errorbar", width = .75) +
  geom_boxplot(
    outlier.size = 0,
    alpha = 0.5,
    coef = 0,
    outlier.shape = NA,
    
  ) +
  stat_summary(
    fun.y = median,
    geom = "smooth",
    position = position_dodge(0.75),
    aes(group = method),
    lwd = 0.8
  ) +
  scb_graphs +
  background_grid("none") +
  geom_vline(
    xintercept = 11.5,
    size = 1.5,
    color = "black",
    linetype = "dashed"
  ) +
  labs(x = "prior error",
       y = "MCC",
       fill = "") +
  theme(
    legend.position = "right",
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    axis.text.x = element_text(
      hjust = 0,
      vjust = 0.5,
      angle = -45,
      size = 12
    ),
    axis.title.x = element_text(size = 14,
                                margin = margin(-1, 0, 0, 0, unit = "lines")),
    plot.margin = margin(0.2, 1, 0.2, 0.2, "cm")
  )
simulation_mcc
```

## Prior completeness

TODO: update once all results/iterations are done.

```{r prior_completeness}
# minimal subset size, prior noise progression
res <- read_tsv("test_result_prior_completeness.txt") %>%
  mutate(fraction = paste0("F=", gsub("fraction", "", fraction_to_keep)))

# create nicer method names
tab <- res %>% 
  mutate(comparison = gsub("bdgraph$", "bdgraph (priors)", comparison),
         comparison = gsub("glasso$", "glasso (priors)", comparison),
         comparison = gsub("bdgraph_no_priors$", "bdgraph (empty)", comparison),
         comparison = gsub("bdgraph_no_priors_full$", "bdgraph (full)", comparison),
         comparison = gsub("glasso_no_priors","glasso", comparison)) %>%
  filter(comparison != "bdgraph (full)") %>%
  dplyr::rename(method=comparison) %>%
  mutate(method = gsub("bdgraph \\(empty\\)", "bdgraph", method)) %>%
  mutate(method = factor(
    method,
    ordered = T,
    levels = c(
      "bdgraph",
      "bdgraph (priors)",
      "glasso",
      "glasso (priors)",
      "irafnet",
      "genie3",
      "genenet"
    )
  ))

# get the MCC plot
simulation_mcc <- ggplot(tab,
                         aes(y = MCC,
                             x = fraction,
                             color = method)) +
  stat_boxplot(geom = "errorbar", width = .75) +
  geom_boxplot(
    outlier.size = 0,
    alpha = 0.5,
    coef = 0,
    outlier.shape = NA,
    
  ) +
  stat_summary(
    fun.y = median,
    geom = "smooth",
    position = position_dodge(0.75),
    aes(group = method),
    lwd = 0.8
  ) +
  scb_graphs +
  background_grid("none") +
  labs(x = "fraction of prior information kept",
       y = "MCC",
       fill = "") +
  theme(
    legend.position = "right",
    legend.text = element_text(size = 12),
    legend.title = element_text(size = 14),
    axis.text.x = element_text(
      hjust = 0,
      vjust = 0.5,
      angle = -45,
      size = 12
    ),
    plot.margin = margin(0.2, 1, 0.2, 0.2, "cm")
  )
simulation_mcc
```
Caption: Analysis of the effect of 'prior completeness' on inference performance.
Y-axis shows the MCC of inferred graphs VS the simulated ground truth, x-axis indicates the fraction
of prior information retained for inference, increasing from left to right. Colors indicate the different models. Boxplots show medians (horizontal line) and first and third quartiles (lower/upper box borders). Whiskers show 1.5*IQR (inter-quartile range).

