---
title: Replication validation based on reviewer comments
author: Johann Hawe | johann.hawe@tum.de
output:
  pdf_document: 
    toc: yes
  html_document: default
---

```{r setup, echo = FALSE, message = FALSE, warning = FALSE}
library(BDgraph)
library(igraph)
library(graph)
library(readr)
library(dplyr)
library(ggplot2)
library(parallel)
library(reshape2)
library(cowplot)
theme_set(theme_cowplot() + background_grid(major="xy"))

source("scripts/validation_methods.R")
source("scripts/lib.R")

threads <- 4

knitr::opts_chunk$set(echo = FALSE)

```

## Load data

We load all previously generated data, including the collected priors and the
model fits for KORA and LOLIPOP.

```{r load_data}
fpriors <- 
  list.files("results/current/biogrid_stringent/priors/", 
             "*.rds", full.names = T)
fkora_data <- 
  list.files("results/current/biogrid_stringent/cohort_data_tfa/kora/", 
             "*.rds", full.names = T)
flolipop_data <- 
  list.files("results/current/biogrid_stringent/cohort_data_tfa/lolipop/", 
             "*.rds", full.names = T)
fkora_fits <- 
  list.files("results/current/biogrid_stringent/fits_tfa/_rerun/kora/", 
             "*.rds", full.names = T)
flolipop_fits <- 
  list.files("results/current/biogrid_stringent/fits_tfa/_rerun/lolipop/", 
             "*.rds", full.names = T)

rename_files <- function(file_list) {
  names(file_list) <- gsub(".rds", "", gsub(".*/", "", file_list))
  file_list
}

fpriors <- rename_files(fpriors)
fkora_data <- rename_files(fkora_data)
flolipop_data <- rename_files(flolipop_data)
fkora_fits <- rename_files(fkora_fits)
flolipop_fits <- rename_files(flolipop_fits)

sentinels <- names(fpriors)
```

- Number of sentinels to be processed: `r length(sentinels)`

## Prior importance for replication

```{r helper_definitions, echo = FALSE}
PSEUDO_PRIOR <- 1e-7

# helper to extract prior edge/non edge sets
get_prior_edge_sets <- function(prior_matrix) {
  
  prior_matrix_tibble <- prior_matrix %>%
    melt() %>% as_tibble()
  
  prior_edges <-  prior_matrix_tibble %>%
    filter(value > PSEUDO_PRIOR) %>%
    dplyr::rename(node1 = Var1, node2 = Var2) %>%
    mutate(node1 = as.character(node1), node2 = as.character(node2)) %>%
    filter(node1 != node2) %>%
    dplyr::select(-value)
  
  non_prior_edges <- prior_matrix_tibble %>%
    filter(value == PSEUDO_PRIOR) %>%
    dplyr::rename(node1 = Var1, node2 = Var2) %>%
    mutate(node1 = as.character(node1), node2 = as.character(node2)) %>%
    filter(node1 != node2) %>%
    dplyr::select(-value)
  return(list(prior_edges = prior_edges, non_prior_edges = non_prior_edges))
}

# helper to get tibble of edges from a graph object
get_edges_from_graph <- function(graph, type = "undirected") {
  graph_nodes <- nodes(graph)
  temp <- t(graph::edgeMatrix(graph)) %>%
    as_tibble() %>%
    mutate(from = graph_nodes[from],
           to = graph_nodes[to]) %>%
    dplyr::rename(node1 = from, node2 = to)
  
  if (type %in% "undirected") {
    # add the 'reverse' edge, too, as we investigated undirected graphs
    bind_rows(temp, dplyr::rename(temp, node1 = node2, node2 = node1))
  } else if(type %in% "directed") {
    sort_nodes(temp)
  } else {
    warning("type not supported.")
  }
}

# helper to determine the set of replicated (common) edges between two graphs (e.g.
# from KORA and LOLIPOP)
get_replicated_edges <- function(graph1, graph2, type = "undirected") {
  graph1_edges <- get_edges_from_graph(graph1, type)
  graph2_edges <- get_edges_from_graph(graph2, type)
  
  replicated_edges <- inner_join(graph1_edges, graph2_edges,
                                 by = c("node1", "node2"))
  replicated_edges
}

# helper to enrich a set of edges over the other
# e.g.: prior edges (set1_edges) are enriched over non_prior edges (set2_edges)
# with respect to replicated edges (background_edges)
# All params are data frames/tibbles with columns "node1" and "node2", both 
# edge directions need to be presented
# as we always have both 'directions' for the undirected edges in the tibbles
# devide measures by 2
get_enrichment_for_edgesets <- function(set1_edges, 
                                        set2_edges,
                                        background_edges) {
  
  set1_in_background <-
    nrow(inner_join(set1_edges, background_edges,
                    by = c("node1", "node2"))) / 2
  
  set1_not_in_background <-
    (nrow(set1_edges) / 2) - set1_in_background
  
  set2_in_background <-
    nrow(inner_join(set2_edges, background_edges,
                    by = c("node1", "node2"))) / 2
  
  set2_not_in_background <-
    (nrow(set2_edges) / 2) - set2_in_background
  
  contingency <- matrix(
    c(
      set1_in_background,
      set1_not_in_background,
      set2_in_background,
      set2_not_in_background
    ),
    ncol = 2,
    byrow = T
  )
  
  return(
    fisher.test(contingency, alternative = "greater") %>%
      broom::tidy() %>%
      mutate(
        set1_in_background = set1_in_background,
        set1_not_in_background = set1_not_in_background,
        set2_in_background = set2_in_background,
        set2_not_in_background = set2_not_in_background
      )
  )
}

# helper to sort nodes in the node columns (node1, node2) so that the lexicographically
# smaller node name is saved in node1 and vice versa
sort_nodes <- function(node_data) {
  node_data %>%
    rowwise() %>%
    mutate(temp1 = sort(c(node1, node2))[1], temp2 =  sort(c(node1, node2))[2],
           node1=temp1, node2=temp2) %>%
    dplyr::select(-temp1, -temp2)
}

get_edge_tibble <- function(graph, threads=1) {
 edges <- graph::edges(graph)
 if(length(unlist(edges)) > 0) {
 parallel::mclapply(names(edges), function(node1) {
   targets <- edges[[node1]]
   
   # ensure sorted by nodename -> enables to remove 'duplicate' edges easily
   lapply(targets, function(node2) {
       nodes <- sort(c(node1, node2))
       tibble(node1=nodes[1], node2=nodes[2])
   }) %>% bind_rows()
   
 }, mc.cores = threads) %>% bind_rows() %>% 
   distinct()
 } else {
   tibble(node1=NA_character_, node2=NA_character_) %>% drop_na()
 }
}

```

We investigate whether available prior information largely drives replication performance.
In brief, we create a contingency table from all sentinels, where we check 1) whether an edge is replicated or not and 2) whether the edge has a prior assigned it or not.

```{r prior_importance_replication}
# iterate over all sentinels and gather stats
result <- mclapply(sentinels, function(sentinel) {
  
  prior_matrix <- readRDS(fpriors[[sentinel]])
  temp <- get_prior_edge_sets(prior_matrix)
  prior_edges <- temp$prior_edges
  non_prior_edges <- temp$non_prior_edges
  
  kora_fits <- readRDS(fkora_fits[[sentinel]])
  lolipop_fits <- readRDS(flolipop_fits[[sentinel]])
  
  graph_models <- names(kora_fits)[!grepl("_fit", names(kora_fits))]
  
  lapply(graph_models, function(graph_model) {
    
    replicated_edges <- get_replicated_edges(kora_fits[[graph_model]],
                                             lolipop_fits[[graph_model]])
    
    get_enrichment_for_edgesets(prior_edges, 
                                non_prior_edges, 
                                replicated_edges) %>%
      mutate(graph_model = graph_model,
             sentinel = sentinel,
             prior_and_replicated = set1_in_background,
             prior_not_replicated = set1_not_in_background,
             not_prior_and_replicated = set2_in_background,
             not_prior_not_replicated = set2_not_in_background) %>% 
      dplyr::select(graph_model, -method, -alternative, everything())
  }) %>% bind_rows()
}, mc.cores = threads) %>% bind_rows() %>%
    mutate(is_prior_based = case_when(graph_model %in% c("glasso", 
                                                         "bdgraph", 
                                                         "irafnet") ~ TRUE,
                                    TRUE ~ FALSE))
result

ggplot(result %>% mutate(estimate = case_when(is.infinite(estimate) ~ 1e4, 
                                              TRUE ~ estimate)), 
       aes(x=log10(estimate), 
           group=is_prior_based, 
           color=is_prior_based)) + 
  geom_freqpoly()
```

Above are the initial 'full' results, i.e. for ea ch graph model and sentinel,
we see the evaluation by use of the fisher test.

The plot indicates the distribution (log10) of estimates over all models and sentinels.
`estimates` amounging to `Inf` were substituted by `10e4` and result in the 
peaks located at the far right of the plot.

Now we look at a simple summary, where we sum up all contingency tables and
calculate a single fisher test for each of the graph models.

```{r prior_importance_replication_summarized}
summarized <- group_by(result, graph_model) %>%
  summarize(prior_and_replicated = sum(prior_and_replicated),
             prior_not_replicated = sum(prior_not_replicated),
             not_prior_and_replicated = sum(not_prior_and_replicated),
             not_prior_not_replicated = sum(not_prior_not_replicated))

summarized_fishertest <- lapply(summarized$graph_model, function(current_model) {
  current_summarized <-
    filter(summarized, graph_model == current_model)
  
  contingency_summarized <-
    matrix(
      c(
        current_summarized$prior_and_replicated,
        current_summarized$prior_not_replicated,
        current_summarized$not_prior_and_replicated,
        current_summarized$not_prior_not_replicated
      ),
      ncol = 2,
      byrow = TRUE
    )
  fisher.test(contingency_summarized, alternative = "greater") %>% 
    broom::tidy() %>%
    mutate(graph_model = current_model) %>%
    dplyr::select(graph_model, everything())
}) %>% bind_rows() %>%
  left_join(summarized, by="graph_model") %>%
  mutate(is_prior_based = case_when(graph_model %in% c("glasso", 
                                                       "bdgraph", 
                                                       "irafnet") ~ TRUE,
                                    TRUE ~ FALSE))

summarized_fishertest

ggplot(summarized_fishertest, aes(y=log10(estimate), 
                                  x=reorder(graph_model, log10(estimate)),
                                  fill = is_prior_based)) + 
  geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle=-45, hjust=0, vjust=0)) + 
  labs(x="graph model")
```

Now we also have a look at a different 'replication': We compare the inferred
network in one cohort to a simple correlation network retrieved from the respective
other cohort. Any replication we see is then not due to any prior information, but rather
solely based on the observed data/likelihoods.

```{r evaluate_correlation_graphs}
fcorrelation_graphs <- 
  list.files("results/current/biogrid_stringent/correlation_graphs_tfa/_rerun/", 
             "*.rds", full.names = T)

fcorrelation_graphs <- rename_files(fcorrelation_graphs)

result_table <- lapply(sentinels, function(sentinel) {
  
  print(sentinel)
 
  correlation_graphs <- readRDS(fcorrelation_graphs[sentinel])
  correlation_graph_kora <- correlation_graphs[["kora"]]$graph
  correlation_graph_lolipop <- correlation_graphs[["lolipop"]]$graph
  
  fit_kora <- readRDS(fkora_fits[sentinel])
  fit_lolipop <- readRDS(flolipop_fits[sentinel])
  
  graph_models <- names(fit_kora)[!grepl("_fit", names(fit_kora))]
  
  lapply(graph_models, function(model) {
    inferred_graph_kora <- fit_kora[[model]]
    inferred_graph_lolipop <- fit_lolipop[[model]]
    
    common_nodes_kora <-
      intersect(nodes(correlation_graph_lolipop),
                nodes(inferred_graph_kora))
    common_nodes_lolipop <-
      intersect(nodes(correlation_graph_kora),
                nodes(inferred_graph_lolipop))
    
    correlation_graph_kora <- subGraph(common_nodes_lolipop,
                                       correlation_graph_kora)
    correlation_graph_lolipop <- subGraph(common_nodes_kora,
                                          correlation_graph_lolipop)
    inferred_graph_kora <- subGraph(common_nodes_kora,
                                    inferred_graph_kora)
    inferred_graph_lolipop <- subGraph(common_nodes_lolipop,
                                       inferred_graph_lolipop)
    
    mcc_kora <- 
      BDgraph::compare(as(inferred_graph_kora, "matrix"),
                       as(correlation_graph_lolipop, "matrix"))["MCC", "estimate1"]
    
    mcc_lolipop <-
      BDgraph::compare(as(inferred_graph_lolipop, "matrix"),
                       as(correlation_graph_kora, "matrix"))["MCC", "estimate1"]
    
    tibble(
      sentinel = sentinel,
      graph_model = model,
      MCC = c(mcc_kora, mcc_lolipop)
    )
    
  }) %>% bind_rows()
}) %>% bind_rows()


result_table %>%
  ggplot(aes(x=reorder(graph_model, -MCC, FUN=median),
             y=MCC)) +
  geom_boxplot(outlier.shape = NA) + 
  geom_point(position=position_jitter(width = 0.15),
             alpha=0.1) +
  #ggpubr::stat_compare_means() +
  labs(title="",
       y="MCC",
       x="method") + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.x = element_blank(),
        axis.text.x = element_text(hjust=0, vjust=0.5, angle=-45, size=12),
        legend.position = "bottom",
        legend.text = element_text(size=14),
        legend.title = element_text(size=15))

  
```

```{r evaluate_correlations}
# takes about 15min with 20 threads available
sink("temp_sentinels.log")
result_table <- mclapply(sentinels, function(sentinel) {
  
  print(sentinel)
 
  correlations <- readRDS(fcorrelation_graphs[sentinel])
  correlations_kora <- sort_nodes(correlations[["kora"]]$fit)
  correlations_lolipop <- sort_nodes(correlations[["lolipop"]]$fit)
  
  fit_kora <- readRDS(fkora_fits[sentinel])
  fit_lolipop <- readRDS(flolipop_fits[sentinel])
  
  prior_matrix <- readRDS(fpriors[[sentinel]])
  temp <- get_prior_edge_sets(prior_matrix)
  prior_edges <- temp$prior_edges %>%
    mutate(is_prior = TRUE)
  
  graph_models <- names(fit_kora)[!grepl("_fit", names(fit_kora))]
  
  lapply(graph_models, function(model) {
    inferred_graph_kora <- fit_kora[[model]]
    inferred_graph_lolipop <- fit_lolipop[[model]]
    
    replicated_edges <- get_replicated_edges(inferred_graph_kora,
                                            inferred_graph_lolipop,
                                            type = "directed") %>%
      mutate(is_replicated = TRUE)
    
    edges_kora <- get_edge_tibble(inferred_graph_kora, 1) %>%
      left_join(y=replicated_edges, by=c("node1" = "node1", "node2" = "node2")) %>%
      left_join(y=prior_edges, by=c("node1" = "node1", "node2" = "node2"))
    edges_lolipop <- get_edge_tibble(inferred_graph_lolipop, 1) %>%
      left_join(y=replicated_edges, by=c("node1" = "node1", "node2" = "node2")) %>%
      left_join(y=prior_edges, by=c("node1" = "node1", "node2" = "node2"))
      
    bind_rows(
      left_join(edges_kora, 
                correlations_kora,
                by = c("node1" = "node1", "node2" = "node2")
      ),
      left_join(edges_lolipop,
                correlations_lolipop,
                by = c("node1", "node2"))
    ) %>%
      mutate(sentinel = sentinel, graph_model = model) %>%
      mutate(is_replicated = case_when(is.na(is_replicated) ~ FALSE,
                                     TRUE ~ TRUE)) %>%
      mutate(is_prior = case_when(is.na(is_prior) ~ FALSE,
                                     TRUE ~ TRUE)) %>%
      mutate(prior_replicated = case_when(is_replicated & is_prior ~ "prior_and_replicated",
                                          is_prior & !is_replicated ~ "only_prior",
                                          is_replicated & !is_prior ~ "only_replicated",
                                          TRUE ~ "none" ))
      
  }) %>% bind_rows()
}, mc.cores = threads) %>% bind_rows()

result_table <- filter(result_table, 
                       
                       graph_model %in% c("glasso", "glasso_no_priors",
                                          "bdgraph", "bdgraph_no_priors"))

summary_table <- group_by(result_table, graph_model, prior_replicated) %>%
  dplyr::summarize(total_edges = n(),
                   cor_upper_quart = quantile(abs(corr), 0.75))

colors <- RColorBrewer::brewer.pal(4, "Paired")

# discuss
category_colors <- c("prior_and_replicated" = colors[2],
                     "only_replicated" = colors[4],
                     "only_prior" = colors[1],
                     "none" = colors[3])

gp <- result_table %>%
  ggplot(aes(x=graph_model,
             y=abs(corr), fill=prior_replicated)) +
  geom_boxplot(outlier.shape = NA) + 
  geom_label(data=summary_table,
             position=position_dodge(1),
             aes(label=total_edges, 
                 y=cor_upper_quart+0.1, 
                 x=graph_model, group=prior_replicated)) +
  scale_fill_manual(values = category_colors) +
  labs(title="",
       y="correlation",
       x="method") + 
  theme(plot.title = element_text(hjust = 0.5),
        axis.title.x = element_blank(),
        axis.text.x = element_text(hjust=0, vjust=0.5, angle=-45, size=12),
        legend.position = "bottom",
        legend.text = element_text(size=14),
        legend.title = element_text(size=15))
gp

```

## Prior importance inferred edge set

Contrary to the replicated edge set above, we can also have a look at the 'inferred' edge set,
which are defined as the union of the edge sets from KORA and LOLIPOP.

```{r prior_importance_inferred_edgeset}
PSEUDO_PRIOR <- 1e-7


# iterate over all sentinels and gather stats
result <- mclapply(sentinels, function(sentinel) {
  
  prior_matrix <- readRDS(fpriors[[sentinel]])
  temp <- get_prior_edge_sets(prior_matrix)
  prior_edges <- temp$prior_edges
  non_prior_edges <- temp$non_prior_edges
  
  kora_fits <- readRDS(fkora_fits[[sentinel]])
  lolipop_fits <- readRDS(flolipop_fits[[sentinel]])
  
  graph_models <- names(kora_fits)[!grepl("_fit", names(kora_fits))]
  
  lapply(graph_models, function(graph_model) {
    kora_edges <- get_edges_from_graph(kora_fits[[graph_model]])
    lolipop_edges <- get_edges_from_graph(lolipop_fits[[graph_model]])
    
    full_edge_set <- bind_rows(kora_edges, lolipop_edges) %>%
      distinct()
    
    get_enrichment_for_edgesets(prior_edges, 
                                non_prior_edges, 
                                full_edge_set) %>%
      mutate(graph_model = graph_model,
             prior_and_inferred = set1_in_background,
             prior_not_inferred = set1_not_in_background,
             not_prior_and_inferred = set2_in_background,
             not_prior_not_inferred = set2_not_in_background) %>% 
      dplyr::select(graph_model, -method, -alternative, everything())
  }) %>% bind_rows() %>%
    mutate(sentinel = sentinel) %>%
    dplyr::select(sentinel, everything())
}, mc.cores = 4) %>% bind_rows() %>%
    mutate(is_prior_based = case_when(graph_model %in% c("glasso", 
                                                       "bdgraph", 
                                                       "irafnet") ~ TRUE,
                                    TRUE ~ FALSE))
result

ggplot(result %>% mutate(estimate = case_when(is.infinite(estimate) ~ 1e4, 
                                              TRUE ~ estimate)), 
       aes(x=log10(estimate), 
           group=is_prior_based, 
           color=is_prior_based)) + 
  geom_freqpoly()

```

```{r prior_importance_inferred_summarized}
summarized <- group_by(result, graph_model) %>%
  summarize(prior_and_inferred = sum(prior_and_inferred),
             prior_not_inferred = sum(prior_not_inferred),
             not_prior_and_inferred = sum(not_prior_and_inferred),
             not_prior_not_inferred = sum(not_prior_not_inferred))

summarized_fishertest <- lapply(summarized$graph_model, function(current_model) {
  current_summarized <-
    filter(summarized, graph_model == current_model)
  
  contingency_summarized <-
    matrix(
      c(
        current_summarized$prior_and_inferred,
        current_summarized$prior_not_inferred,
        current_summarized$not_prior_and_inferred,
        current_summarized$not_prior_not_inferred
      ),
      ncol = 2,
      byrow = TRUE
    )
  fisher.test(contingency_summarized, alternative = "greater") %>% 
    broom::tidy() %>%
    mutate(graph_model = current_model) %>%
    dplyr::select(graph_model, everything())
}) %>% bind_rows() %>%
  left_join(summarized, by="graph_model") %>%
  mutate(is_prior_based = case_when(graph_model %in% c("glasso", 
                                                       "bdgraph", 
                                                       "irafnet") ~ TRUE,
                                    TRUE ~ FALSE))

summarized_fishertest

ggplot(summarized_fishertest, aes(y=log10(estimate), 
                                  x=reorder(graph_model, log10(estimate)),
                                  fill = is_prior_based)) + 
  geom_bar(stat="identity") + 
  theme(axis.text.x = element_text(angle=-45, hjust=0, vjust=0)) + 
  labs(x="graph model")
```

## Replication with prior noise

In a next step, we investigate replication performance under consideration of 
differing levels of priors noise.
First we have a look at the enrichment of prior edges over non-prior edges within the set of replicated edges.

> NOTE: at the beginning we defined the edges sets using the ORIGINAL (not noisified) prior matrix. We now switched to using the noisified one. Naturally, this changes our previous obesrvation of decreased enrichment with increase in prior noise (we rather observed only the 'decrease in true signal in the priors' than any performance changes...)

```{r load_data_prior_noise}

base_dir <- "results/current/biogrid_stringent/"
fit_files <- list.files(paste0(base_dir, "fits_tfa/replication_prior_noise/"), 
                        "*.rds", full.names = T)
names(fit_files) <-  gsub(".rds","", basename(fit_files))

# we need the original priors to define true/false prior edges in noisy data
prior_files <- list.files(paste0(base_dir, "priors/"), 
                          ".*.rds", full.names = T)
names(prior_files) <- gsub(".rds","", basename(prior_files))

models <- c("bdgraph", "bdgraph_no_priors", "glasso", "glasso_no_priors",
            "genenet", "genie3", "irafnet")

result_table <- parallel::mclapply(names(fit_files), function(locus) {
  
  file <- fit_files[locus]
  sentinel <- gsub("_meqtl|_eqtlgen", "", locus)
  locus_type <- gsub("rs.*_", "", locus)
  
  fits <- readRDS(file)
  
  lapply(names(fits), function(noise_level) {
    
    fits_at_noise_level <- fits[[noise_level]]
    
    fits_kora <- fits_at_noise_level[["kora"]]
    fits_lolipop <- fits_at_noise_level[["lolipop"]]
    
    priors <- fits_at_noise_level$priors
    
    prior_edge_sets <- get_prior_edge_sets(priors)
    prior_edges <- prior_edge_sets$prior_edges
    non_prior_edges <- prior_edge_sets$non_prior_edges
    
    lapply(models, function(model) {
      fit_kora <- fits_kora[[model]]
      fit_lolipop <- fits_lolipop[[model]]
      
      replicated_edges <- get_replicated_edges(fit_kora, fit_lolipop)
    
    get_enrichment_for_edgesets(prior_edges, 
                                non_prior_edges, 
                                replicated_edges) %>%
        mutate(graph_model = model, 
               sentinel = sentinel, 
               locus_type = locus_type,
               noise_level = gsub("noise_level_", "", noise_level),
               prior_and_replicated = set1_in_background,
               prior_not_replicated = set1_not_in_background,
               not_prior_and_replicated = set2_in_background,
               not_prior_not_replicated = set2_not_in_background) %>% 
        dplyr::select(sentinel, locus_type, graph_model, -method, -alternative, 
                      everything())
      
    }) %>% bind_rows()
    
  }) %>% bind_rows()
  
}, mc.cores = threads) %>% bind_rows() %>%
   mutate(is_prior_based = case_when(graph_model %in% c("glasso", 
                                                         "bdgraph", 
                                                         "irafnet") ~ TRUE,
                                    TRUE ~ FALSE)) %>%
  mutate(graph_model = gsub("^glasso$", "glasso (priors)", graph_model)) %>%
  mutate(graph_model = gsub("^bdgraph$", "bdgraph (priors)", graph_model)) %>%
  mutate(graph_model = gsub("_no_priors", "", graph_model))

# define plot colors
paired <- RColorBrewer::brewer.pal(4, "Paired")
names(paired) <- c("glasso", "glasso (priors)", "bdgraph", "bdgraph (priors)")
unpaired <- RColorBrewer::brewer.pal(7, "Dark2")[c(2,3,7)]
names(unpaired) <- c("irafnet", "genie3", "genenet")
graph_cols <- c(paired, unpaired)

max_non_infinite_estimate <- filter(result_table, is.finite(estimate)) %>%
  filter(estimate == max(estimate)) %>% pull(estimate)

gp1 <- ggplot(
  result_table %>%
    mutate(
      estimate = case_when(
        is.infinite(estimate) ~ max_non_infinite_estimate,
        TRUE ~ estimate
      )
    ),
  aes(
    y = log10(estimate),
    x = noise_level,
    color = reorder(graph_model, log10(estimate), mean)
  )
) +
  geom_boxplot(outlier.shape = NA) +
  stat_summary(
    fun.y = median,
    geom = "smooth",
    position = position_dodge2(width = 0.9),
    aes(group = reorder(graph_model, log10(estimate), mean)),
    lwd = 0.8
  ) +
  geom_hline(
    yintercept = 0,
    color = "grey",
    linetype = "dashed",
    size = 1
  ) +
  scale_color_manual(values = graph_cols) +
  labs(color = "model",
       subtitle = "Enrichment of prior over non-prior edges\nin replicated edges")

gp1

```

Now we check out the enrichment of **false prior** edges over the **true prior** edges within the set of replicated edges.

```{r enrich_false_over_true_prior_edges}
get_false_prior_edges <- function(noisy_priors, original_priors) {
  noisy_sets <- get_prior_edge_sets(noisy_priors)
  original_sets <- get_prior_edge_sets(original_priors)
  
  # false priors: priors defined in noisy set which are part of the non-prior set
  # in the original data
  semi_join(noisy_sets$prior_edges, original_sets$non_prior_edges,
            by=c("node1", "node2"))
}

get_true_prior_edges <- function(noisy_priors, original_priors) {
  noisy_sets <- get_prior_edge_sets(noisy_priors)
  original_sets <- get_prior_edge_sets(original_priors)
  
  # true priors: priors defined in noisy set which are part of the prior set
  # in the original data
  semi_join(noisy_sets$prior_edges, original_sets$prior_edges,
            by=c("node1", "node2"))
}

result_table <- parallel::mclapply(names(fit_files), function(locus) {
  
  file <- fit_files[locus]
  sentinel <- gsub("_meqtl|_eqtlgen", "", locus)
  locus_type <- gsub("rs.*_", "", locus)
  
  fits <- readRDS(file)
  
  priors <- readRDS(prior_files[locus])
  
  lapply(names(fits), function(noise_level) {
    
    if(noise_level == "noise_level_0") return(NULL)
    
    fits_at_noise_level <- fits[[noise_level]]
    
    fits_kora <- fits_at_noise_level[["kora"]]
    fits_lolipop <- fits_at_noise_level[["lolipop"]]
    
    # the (noisy) priors used in the inference analysis
    noisy_priors <- fits_at_noise_level$priors
    false_prior_edges <- get_false_prior_edges(noisy_priors, priors)
    true_prior_edges <- get_true_prior_edges(noisy_priors, priors)
    
    lapply(models, function(model) {
      fit_kora <- fits_kora[[model]]
      fit_lolipop <- fits_lolipop[[model]]
      
      replicated_edges <- get_replicated_edges(fit_kora, fit_lolipop)
    
      false_over_true_prior <-
        get_enrichment_for_edgesets(false_prior_edges,
                                    true_prior_edges,
                                    replicated_edges) %>%
        mutate(
          graph_model = model,
          sentinel = sentinel,
          locus_type = locus_type,
          noise_level = gsub("noise_level_", "", noise_level)) %>%
        dplyr::select(sentinel,
                      locus_type,
                      graph_model,-method,-alternative,
                      everything())
      
      false_over_true_prior
      
    }) %>% bind_rows()
    
  }) %>% bind_rows()
  
}, mc.cores = threads) %>% bind_rows() %>%
   mutate(is_prior_based = case_when(graph_model %in% c("glasso", 
                                                         "bdgraph", 
                                                         "irafnet") ~ TRUE,
                                    TRUE ~ FALSE)) %>%
  mutate(graph_model = gsub("^glasso$", "glasso (priors)", graph_model)) %>%
  mutate(graph_model = gsub("^bdgraph$", "bdgraph (priors)", graph_model)) %>%
  mutate(graph_model = gsub("_no_priors", "", graph_model))

max_non_infinite_estimate <- filter(result_table, is.finite(estimate)) %>%
  filter(estimate == max(estimate)) %>% pull(estimate) %>% unique

result_table <- result_table %>%
  mutate(estimate = case_when(
    is.infinite(estimate) ~ max_non_infinite_estimate,
    TRUE ~ estimate
  ))


gp2 <- ggplot(result_table,
              aes(
                y = log10(estimate),
                x = noise_level,
                color = reorder(graph_model, log10(estimate), mean)
              )) +
  geom_boxplot(outlier.shape = NA) +
  stat_summary(
    fun.y = median,
    geom = "smooth",
    position = position_dodge2(width = 0.9),
    aes(group = reorder(graph_model, log10(estimate), mean)),
    lwd = 0.8
  ) +
  geom_hline(
    yintercept = 0,
    color = "grey",
    linetype = "dashed",
    size = 1
  ) +
  scale_color_manual(values = graph_cols) +
  labs(color = "model",
       subtitle = "Enrichment of false priors over true priors\nin replicated edges.")

ggpubr::ggarrange(
  gp1,
  gp2,
  ncol = 2,
  labels = "AUTO",
  common.legend = T,
  legend = "bottom"
)
```

Above plot shows that edges (false/true) are treated the same of different scenarios (?).

```{r}

result_table <- parallel::mclapply(names(fit_files), function(locus) {
  
  file <- fit_files[locus]
  sentinel <- gsub("_meqtl|_eqtlgen", "", locus)
  locus_type <- gsub("rs.*_", "", locus)
  
  fits <- readRDS(file)
  
  priors <- readRDS(prior_files[locus])
  
  lapply(names(fits), function(noise_level) {
    
    if(noise_level == "noise_level_0") return(NULL)
    
    fits_at_noise_level <- fits[[noise_level]]
    
    fits_kora <- fits_at_noise_level[["kora"]]
    fits_lolipop <- fits_at_noise_level[["lolipop"]]
    
    # the (noisy) priors used in the inference analysis
    noisy_priors <- fits_at_noise_level$priors
    
    false_prior_edges <- get_false_prior_edges(noisy_priors, priors)
    
    noisy_edge_sets <- get_prior_edge_sets(noisy_priors)
    noisy_non_prior_edges <- noisy_edge_sets$non_prior_edges
    
    lapply(models, function(model) {
      fit_kora <- fits_kora[[model]]
      fit_lolipop <- fits_lolipop[[model]]
      
      replicated_edges <- get_replicated_edges(fit_kora, fit_lolipop)
    
      false_over_non_prior <-
        get_enrichment_for_edgesets(false_prior_edges,
                                    noisy_non_prior_edges,
                                    replicated_edges) %>%
        mutate(
          graph_model = model,
          sentinel = sentinel,
          locus_type = locus_type,
          noise_level = gsub("noise_level_", "", noise_level)) %>%
        dplyr::select(sentinel,
                      locus_type,
                      graph_model,-method,-alternative,
                      everything())
      
      false_over_non_prior
      
    }) %>% bind_rows()
    
  }) %>% bind_rows()
  
}, mc.cores = threads) %>% bind_rows() %>%
   mutate(is_prior_based = case_when(graph_model %in% c("glasso", 
                                                         "bdgraph", 
                                                         "irafnet") ~ TRUE,
                                    TRUE ~ FALSE)) %>%
  mutate(graph_model = gsub("^glasso$", "glasso (priors)", graph_model)) %>%
  mutate(graph_model = gsub("^bdgraph$", "bdgraph (priors)", graph_model)) %>%
  mutate(graph_model = gsub("_no_priors", "", graph_model))

max_non_infinite_estimate <- filter(result_table, is.finite(estimate)) %>%
  filter(estimate == max(estimate)) %>% pull(estimate) %>% unique

result_table <- result_table %>%
  mutate(estimate = case_when(
    is.infinite(estimate) ~ max_non_infinite_estimate,
    TRUE ~ estimate
  ))

gp3 <- ggplot(result_table,
              aes(
                y = log10(estimate),
                x = noise_level,
                color = reorder(graph_model, log10(estimate), mean)
              )) +
  geom_boxplot(outlier.shape = NA) +
  stat_summary(
    fun.y = median,
    geom = "smooth",
    position = position_dodge2(width = 0.9),
    aes(group = reorder(graph_model, log10(estimate), mean)),
    lwd = 0.8
  ) +
  geom_hline(
    yintercept = 0,
    color = "grey",
    linetype = "dashed",
    size = 1
  ) +
  scale_color_manual(values = graph_cols) +
  labs(color = "model",
       subtitle = "Enrichment of false priors over non priors\nin replicated edges.")


```

## Session Info
```{r session_info}
devtools::session_info()
```
