---
title: "Simulation summary"
author: "Johann Hawe"
date: "`r format(Sys.time(), '%d %B, %Y')`"
params:
   rmd: "summarize-simulation.Rmd"
output:
  html_document:
  highlight: tango
  number_sections: no
  theme: default
  toc: yes
  toc_depth: 3
  toc_float:
    collapsed: no
    smooth_scroll: yes
---

```{r}
library(ggplot2)
```

This document summarizes the simmulation study for the bayesian Gaussian Graphical Modelling approach.
For each of the sentinels which had a 'true graph' available (currently based on the network-randwomwalk
approach) we generated data and applied the bdgraph algorithm.

```{r}
# get all results into a single large table
indir <- snakemake@params$indir
#indir <- "../results/current/simulation/validation/"
print(indir)
files <- list.files(indir, "*.txt", full.names = T)
print("Number of files being processed:")
print(length(files))
tab <- c()
for(i in 1:length(files)) {
  f <- read.table(files[i], sep="\t", header=T, row.names = 1)
  
  # we dont need the comparison to the true graph for now
  f <- f[f$comparison != "true",]
  f$R <- paste0("R=", f$rdegree)
  if(i > 1) {
    f <- f[,colnames(tab)]
  }
  tab <- rbind(tab, f)
  colnames(tab) <- colnames(f)
}

```

Overall we are looking at `r nrow(tab)/2` sentinels for we which we have calculated GGM fits with and without prior information, depicted in the plots below as 'ggm.type'. We show summary plots for the individual ggm.types for some standard performance measures (specificity, sensitivity, F1.score) as well as he Matthew's Correlation Coefficient (MCC) which is a correlation measure which can be applied if the individual classes differe largely in size.

```{r}
# jitter the points shown in the boxplot
gp <- geom_point(position = position_jitter(width = 0.15, height = 0))
gb <- geom_boxplot(outlier.shape = NA)
fg <- facet_grid(. ~ R)
fw <- facet_wrap( ~ R, nrow = 2)
theme_update(plot.title = element_text(hjust = 0.5), 
             plot.subtitle = element_text(hjust = 0.5))
theme_update(axis.text.x = element_blank())
# number of samples
N <- length(unique(tab$snp))

pdf("../results/current/simulation/simulation_validation.pdf", width=9, height=9)
ggplot(tab, aes(y=specificity,x=comparison,fill=comparison)) + gb + gp + fw + ggtitle("Specificity", subtitle = paste0("(N = ", N, ")"))
ggplot(tab, aes(y=sensitivity,x=comparison,fill=comparison)) + gb + gp + fw +ggtitle("Sensitivity", subtitle = paste0("(N = ", N, ")"))
ggplot(tab, aes(y=F1.score,x=comparison,fill=comparison)) + gb + gp + fw + 
  ggtitle("F1.score", subtitle = paste0("(N = ", N, ")"))
ggplot(tab, aes(y=MCC,x=comparison,fill=comparison)) + gb + gp + fw +
  ggtitle("MCC", subtitle = paste0("(N = ", N, ")"))
dev.off()

```

<a download="summarize-simulation.Rmd" href="`r base64enc::dataURI(file = params$rmd, mime = 'text/rmd', encoding = 'base64')`">R Markdown source file (to produce this document)</a>
